### 2012-02-23 intention motor representation with corrado sinigaglia/joint action version/intention and motor representation joint action version.tex
started on talk

\def \papersize {a4paper}

@@

\def \ititle {Intention and motor representation:}
\def \isubtitle {how to resist temptation}
\def \iauthor {S. Butterfill* \& C. Sinigaglia**
\\ 
**Department of Philosophy, University of Warwick
\\ 
***Dipartimento di Filosofia, Università degli Studi di Milano}
\def \iemail{s.butterfill@warwick.ac.uk}
%\date{}

@@

\begin{abstract}
Are there distinctive roles for intention and motor representation in explaining the purposiveness of action? Standard accounts of action assign a role to intention but are silent on motor representation. The temptation is to suppose that nothing need be said here because motor representation is either only an enabling condition for purposive action or else merely a variety of intention. This paper provides reasons for resisting temptation. Some motor representations, like intentions, coordinate actions in virtue of representing outcomes; but, unlike intentions, motor representations cannot feature as premises or conclusions in practical reasoning.  This implies that motor representation has a distinctive role in explaining the purposiveness of action. It also gives rise to a problem: were the roles of intention and motor representation entirely independent, this would impair effective action. Resisting temptation therefore requires explaining how intentions interlock with motor representations. The solution, we argue, is to recognise that the contents of intention can be partially determined by the contents of motor representations. Understanding this content-determining relation enables better understanding how intentions are related to actions.
\end{abstract}

@@

What is the relation between a purposive action and the outcome or outcomes to which it is directed? The standard way of answering this question appeals to intention, a propositional attitude which plays a characteristic role in planning and coordinating action, is linked to practical reasoning and is subject to characteristic norms (\citealp{Bratman:1987xw}).  On the standard view, an action is directed to an outcome in virtue of the action's being appropriately related to an intention which represents this outcome or some related outcome. As this view is usually expounded, the relation between actions and outcomes to which they are directed is treated as largely independent of the motor processes and representations underpinning action execution. Motor representations are usually considered either as philosophically irrelevant enabling conditions, or else as merely filling in additional details of the basic schema provided by the standard story. Our aim is to show that this is a mistake. We shall argue that, once some basic features of motor representations are properly understood, the standard view must be  refined and extended in ways that allow us to recognize distinctive roles for intention and motor representation in explaining the purposiveness of action.

Twin temptations stand in our way. The first temptation is to suppose that motor representation, although significant  for action execution, has no role in grounding the outcome-directedness of action. Surrendering to this temptation might be reasonable if all motor representations were about merely kinematic or dynamic features of actions. But some motor representations do represent action outcomes and some actions are directed to those outcomes in virtue of the guiding role of motor representations, or so we shall argue. This reason to resist the first temptation throws us straight into the arms of a second temptation, the temptation to suppose that where motor representations represent action outcomes, they are intentions. Surrendering to this second temptation would be reasonable if motor representations were like intentions in that they could feature as premises or conclusions in practical reasoning, or if there were some other planning processes in which both intention and motor representation could feature. However, as we shall argue, this is impossible due to differences between the propositional format of intentions and the distinctively motor, non-propositional format of motor representations.  In short, we must resist both temptations because some motor representations are like intentions in representing action outcomes while also remaining sufficiently unlike intentions in that no single planning process can integrate both intention and motor representation.

While there are good reasons to resist both temptations (or so we shall argue), the recognition that motor representation can ground the outcome-directedness of purposive actions independently of intention gives rise to a problem. For a single action, which outcomes it is directed to may be multiply determined by an intention and, seemingly independently, by a motor representation. Unless such intentions and motor representations are to pull an agent in incompatible directions, which would typically impair action execution, there are requirements concerning how the outcomes they represent must be related to each other. The problem is to explain how any such requirements could be met non-accidentally; we call this `the interface problem' . The key to solving this problem, we shall suggest, is to recognise that intentions can have constituents which refer to outcomes by deferring to motor representations of those outcomes. Effective action sometimes requires that the contents of intentions and of motor representations interlock, and this interlocking occurs when the contents of intentions are determined in part by the contents of motor representations.  So whereas discussions of intention frequently ignore motor aspects of action, it turns out that understanding how intentions interface with motor representations is required for fully understanding how intentions are related to actions.

@@

Our overall aim is to show that intention and motor representation play distinctive roles in explaining the purposiveness of action, that is, in relating actions to outcomes to which they are directed. The first temptation standing in our way is the temptation to suppose that motor representation, although perhaps important for action execution in some agents, has no bearing at all on our question about the relation between actions and the outcomes to which they are directed. Just as it would be an error to suppose that details of musculoskeletal structure are relevant to this question, so equally---so the temptation---it would be an error to suppose that facts about motor representation are relevant here.

Surrendering to this temptation might be reasonable if all motor representations represented only kinematic or dynamic features of actions, such as mere joint displacements or muscle contractions. However, we shall argue in this section that some motor representations represent action outcomes such as grasping, tearing or throwing. Furthermore, as we shall go on to argue (also in this section), such representations ground purposive actions. This in outline is why the first temptation should be resisted.

How does representing an outcome differ from representing merely kinematic or dynamic features of action? How, for instance, does a motor representation of grasping (assuming for now that such a thing exists) differ from a representation of a sequence of joint displacements? First, a motor representation of grasping captures something common to many different sequences of joint displacements and postures involving a variety of effectors. To illustrate, an agent might grasp an object with her hand, with her mouth, with normal pliers (where grasping requires closing the hand), or with reverse pliers (where grasping requires opening the hand). A motor representation of grasping identifies something common to all these cases. Second, and conversely, a motor representation of grasping potentially distinguishes between the same sequence of joint displacements in different contexts. For instance, how a grasping action is represented may depend on a relatively distal outcome: on whether (say) it results in the object grasped being eaten or placed. Also the joint displacements which realise grasping in one context might in another context realise a different action, such as scratching; or they might fail to realise any action at all (because there is no target object, say). Third, a motor representation of an action outcome locates the action within a hierarchical structure. To illustrate, a left-handed precision grip and a right-handed whole hand grip are different action outcomes and both instances of grasping.\footnote{ On the notion of action outcome in motor representation, see \citet{jeannerod_motor_2006} and \citet{rizzolatti_mirrors_2008}.}

Why accept that there are motor representations of action outcomes? The first step is to consider evidence that motor processes carry information about action outcomes. For any given marker of motor processing (such as a pattern of neuronal discharge or motor-evoked potentials), how can we test whether that marker carries information about action outcomes? The basic principle is straightforward: vary kinematic and dynamic features while holding constant an action outcome; and, conversely, vary action outcomes while holding kinematic and dynamic features constant. In practice researchers have devised many ingenious ways to achieve this. In order to vary kinematic and dynamic features while holding action outcomes constant, in some studies a single action outcome is achieved using different effectors, hand, mouth or foot, say \citep{rizzolatti:1988_functional,Rizzolatti:2001ug,cattaneo:2010_state-dependent}. A variation on this approach is to contrast performing a grasping action with different tools, so that the same action outcome might require closing or opening the hand depending on the tool used \citep{umilta:2008pliers,cattaneo:2009_representation,rochat:2010_responses}. In order to vary action outcome while holding kinematic and dynamic features constant, researchers have contrasted grasping movements with different distal outcomes such as eating and placing \citep{Fogassi:2005nf,bonini:2010_ventral,cattaneo:2007_impairment}. Another approach is to contrast the same grasping movements performed in the presence or manifest absence of a target object \citep{Umilta:2001zr,villiger:2010_activity}. A related alternative is to contrast the same grasping movements in the presence of objects which could, or manifestly could not, be grasped by means of such movements \citep{koch:2010_resonance}. In each of these cases there is evidence that some markers of motor processing are correlated with action outcomes rather than narrowly kinematic or dynamic features of action.\footnote{ Of course some researchers have raised doubts concerning some of the evidence that there are markers of motor processing which carry information about action outcomes (e.g.\ \citealp{cavallo_grasping_2011,borroni_mirroring_2011}).  On balance, however, the evidence supports this conclusion.}

To say that motor processing involves information about action outcomes is not, of course, to say that there are motor representations of action outcomes.  To make the step from information to representation we have to show that information about action outcomes guides processing \citep[compare][]{Dretske:1988sq}. To this end we shall now consider how information about action outcomes is relevant in motor planning and monitoring, which are two of the functional roles of motor representation.

First take motor planning. This involves satisfying a variety of requirements. For example, in grasping a mug it is necessary for the hand to prefigure the shape of the mug, to move towards it avoiding potential obstacles and to reach it at a velocity that is both compatible with achieving the type of grip to be used and also suitable given features of the mug such as its fragility and weight \citep{Jeannerod:1995bb,jeannerod:1998nbo}. The need to plan sequences of actions, which may overlap, imposes further requirements. How one should reach for and then grasp a heavy frying pan (say) depends on what one will then do with it. One way of grasping it might be ideal for safely transporting its contents, another for emptying it. The many requirements on motor planning cannot normally be met by explicit practical reasoning, especially given the rapid and fluid transitions involved in many action sequences. Rather they require motor processes and representations.

Second, motor representations function in monitoring action. They provide inputs to internal predictive models that estimate likely effects of actions. Sensory feedback provides information about the actual course of the action which can be compared against the predictions. Adjustments are made in order to minimise the discrepancy between these \citep{wolpert:1995internal,miall:1996_forward}. Monitoring and control of this kind is a second functional role of motor representation.

These facts about the functional roles of motor representation in planning and monitoring action reveal that it would be advantageous for some such representations to represent action outcomes rather than merely patterns of joint displacement or muscle activation. In the case of planning, many of the requirements to be satisfied are partially dependent on action outcomes and not only on more narrowly kinematic or dynamic features of action. Representation of action outcomes could therefore play a role in simplifying planning processes.  Equally, monitoring involves predicting relatively distal effects of bodily movements including facts about the locations of objects. To make such predictions exclusively in terms of individual joint displacements or muscle activations would be computationally demanding as the human hand alone has over 20 degrees of freedom. Efficient prediction plausibly demands several higher-level representations of action including representations of action outcomes \citep{arbib:1985_coordinated,mason:2001_hand,santello:2002_patterns}. In short, it is clear that motor representations of action outcomes could be useful: in planning, because some requirements for efficient motor planning concern outcomes to which actions are directed; and in monitoring, because representations of outcomes could simplify prediction. Given that, as we saw, markers of motor processes carry information about action outcomes, it is reasonable to conclude from the relevance of action outcomes to the functional roles of motor representation that some motor representations are representations of action outcomes.

So far we have argued that motor processes involve representations of action outcomes. It remains to show that such representations ground purposive actions. But this is hardly a further step. How do intentions ground the purposiveness of actions?  On any standard view, an intention represents an outcome, causes an action, and does so in such a way as to facilitate the outcome's occurrence. Similarly, motor representations of outcomes represent action outcomes, play a role in generating actions, and do this in a way that normally facilitates the occurrence of the outcomes represented. To say that motor representations do all this is one way of making precise the metaphor involved in saying that purposive actions are directed to outcomes.  Moreover, there is a clear resemblance between the natural way of understanding intentions as grounding outcome-directedness and the way in which motor representations ground outcome-directedness.  So given the functional roles and representational objects of motor representations, there is little room for doubt that motor representations can ground the outcome-directedness of purposive actions.

To sum up, the first temptation was to suppose that motor representation has no bearing at all on our question about the relation between actions and the outcomes to which they are directed. One reason to resist this temptation is the fact that, as we have just argued, motor representations not only represent action outcomes but also ground the directedness of actions to outcomes.  This argument might be taken as grounds for giving in to a second temptation.

@@

The second temptation is to suppose that motor representation, or some species of it, is a variety of intention, where intention is understood in the standard way as a propositional attitude with a characteristic role in practical reasoning \citep{Bratman:1987xw}. In this section we explain why this temptation should also be resisted.

As background we first need a generic distinction between content and format. Imagine you are in an unfamiliar city and are trying to get to the central station. A stranger offers you two routes. Each route could be represented by a distinct line on a paper map. The difference between the two lines is a difference in content. Each of the routes could alternatively have been represented by a distinct series of instructions written on the same piece of paper; these cartographic and propositional representations differ in format. The format of a representation constrains its possible contents. For example, a representation with a cartographic format cannot represent what is represented by sentences such as `There could not be a mountain whose summit is inaccessible.'\footnote{ Note that the distinction between content and format is orthogonal to issues about representational medium. The maps in our illustration may be paper map or electronic maps, and the instructions may be spoken, signed or written. This difference is one of medium.} The distinction between content and format is necessary because, as our illustration shows, each can be varied independently of the other.

Our aim in this section is to show that motor representations differ from intentions with respect to their format. This is why the second temptation must be resisted. That motor representations differ in format from intentions shows that they are genuinely distinct phenomena.\footnote{ Readers already convinced that motor representation differs from intention in being non-conceptual will not need the following argument in order to conclude that they are distinct phenomena. However the following considerations also show that motor and perceptual representations differ in format, which will be relevant in later when we consider how motor representations and intentions jointly lead to action.}

How can we distinguish representational formats? Because a representational format is typically associated with characteristic processing profiles, it is sometimes possible to infer similarities and differences in representational format from similarities and differences in the processes in which representations feature. This works both for artifactual and mental representations.  To illustrate in terms of our earlier example, suppose that two people have representations of the same route but for one person the route is represented by a line on a map (so in a cartographic format) whereas the other person has a propositional representation of the route. Some transformations are likely to be easier for the person with the cartographic route representation (depending on the projection used, of course); examples include reversing the route, determining how many times a certain river is crossed and transforming the route into a sequence of compass bearings. Other transformations, such as turning the route into a list of salient landmarks, may be easier for the person with the propositional route description. So some patterns of difference in the two people's performances may be explained by the difference in the formats of their representations. If we did not already know that the two people's route representations involved different formats, we might infer this from the facility with which each performed various transformations of the route.

Cognitive neuroscience frequently depends on inferences of just this type. Contrast imagining seeing an object moving with imagining moving an object. There is evidence that imagining seeing and imagining acting involve different processes. For instance, each can be selectively impaired \citep{sirigu:2011_motor}; and factors such as limb amputation or hand posture can selectively interfere with imagining performing an action \citep{nico:2004_left,vargas:2004s_influence,fourkas:2006_influence}. Importantly for our purposes, imagining seeing an object move and imagining moving an object also differ in ways that indicate differences in representational format. For instance, the way imagining acting unfolds in time is quite different from the way imagining seeing unfolds: how long the former takes is closely related to the time it would take to perform the action, but this is not true of the latter \citep{decety:1989_timing,decety:1996_imagined,Jeannerod:1994oz}. In addition, for object-related actions such as grasping the handle of a cup, manipulating the object in ways that make the action harder (such as orienting the handle to make it less convenient for you to grasp) make a corresponding difference to the effort involved in imagining performing the action \citep{parsons:1994_temporal,frak:2001_orientation}, whereas such manipulations would make no systematic difference to the effort involving in imaging seeing. The explanation for these differences between imagining acting and imagining seeing plausibly involves a difference in the formats of the representations involved.

How do these findings bear on motor representation? Contrast imagining an action with actually performing that action. There is substantial evidence that imagining an action involves many of the same processes which would be involved in performing that action up to the actual muscle contractions \citep{jeannerod:1995_mental,jeannerod:2003_mechanism}. There is also independent evidence that the representations underpinning imagining acting have the same format as those involved in actually acting. This evidence takes the form of similarities in processing profiles. Just as some similarities between imagining seeing something and actually seeing something can be explained on the hypothesis that the representations involved in imagining seeing and actually seeing have a common format \citep{kosslyn:1994_image}, so also some similarities between imagining grasping a cup and actually grasping it are evidence that the representations involved in imagining grasping are of the same format as those involved in actually grasping the cup. We can conclude, then, that imagining and executing actions are both underpinned by motor representations, and that motor representations differ in format from representations involved in both seeing and imagining seeing.

So far we have been arguing that motor and visual representations differ in format.  Why suppose that motor representations also differ in format from intentions?  As we noted earlier, the format of a representation constrains its possible contents. Just as a representation's having a cartographic format restricts its possible contents, so the motor format would imply restrictions on representations having it. In imagining executing actions such as grasping a cup, there are bio-mechanical, dynamical and postural constraints, among others, on what can be imagined. These constraints are closely related to those which govern actually performing such actions \citep{Jeannerod:2001yb} . Such constraints are plausibly explained by the representational format of motor representations. Since no such constraints would be expected to apply where propositional representations are involved, we can conclude (as is widely assumed) that the format of motor representations is not propositional.

To conclude that intention and motor representation are genuinely distinct phenomena it is not quite enough to know that motor representations are non-propositional, of course. In addition we must know that intentions are propositional. We take this claim to be a consequence of the role of intention in practical reasoning and of the fact that one can have intentions involving quantification and identity; for example, one can intend that one cross seven distinct bridges in 48 hours without yet specifying which bridges or hours.\footnote{ Of course some use the term `intention' for non-propositional representations involved in the execution and control of action.  This is a narrowly terminological issue.}

@@

We have just argued for three claims. First, some motor representations represent outcomes (rather than, say, only bodily movements). Second, there are actions whose directedness to an outcome is grounded in motor representation. And third, motor representation differs from intention with respect to representational format. A consequence of these claims is that a single purposive action may involve representations of the outcomes to which it is directed in at least two different representational formats, motor and propositional. This contributes to a problem we call \emph{the interface problem}; in this section we explain how it arises.

Imagine that you are strapped to a spinning wheel facing near certain death as it plunges you into freezing water. To your right you can see a lever and to your left there is a button. In deciding that pulling the lever offers you a better chance of survival than pushing the button, you form an intention to pull the lever, hoping that this will stop the wheel. If things go well, and if intentions are not mere epiphenomena, this intention will result in your reaching for, grasping and pulling the lever. These actions---reaching, grasping and pulling---may be directed to specific outcomes in virtue of motor representations which guide their execution. It shouldn't be an accident that, in your situation, you both intend to pull a lever and you end up with motor representations of reaching for, grasping and pulling that very lever, so that the outcomes specified by your intention match those specified by motor representations. If this match between outcomes variously specified by intentions and by motor representations is not to be accidental, what could explain it?

The interface problem is the problem of answering this question, of explaining how intentions and motor representations, with their distinct representational formats, are related in such a way that, in at least some cases, the outcomes they specify non-accidentally match. But why think that this question poses a problem at all?

Let us start by putting the question more precisely.  First we should define the relevant notion of matching. Two collections of outcomes, A and B, \emph{match} in a particular context just if, in that context, either the occurrence of the A-outcomes would normally constitute or cause, at least partially, the occurrence of the B-outcomes or vice versa. To illustrate, one way of matching is for the B-outcomes to be the A-outcomes. Another way of matching is for the B-outcomes to stand to the A-outcomes as elements of a more detailed plan stand to those of a less detailed one.

Now we can put the question more generally. There are cases in which a particular action is guided both by one or more intentions and by one or more motor representations. In at least some such cases, the outcomes specified by the intentions match the outcomes specified by the motor representations. Furthermore, this match is not always accidental. How do non-accidental matches come about?

In principle one might try to explain the match by supposing that intentions and motor representations have a common cause. If the mere presence (or the mere perception) of a lever invariably triggered intentions and motor representations specifying grasping (say), it might be possible to explain matching in this way. But this sort of consideration cannot provide a full explanation of non-accidental matching for two reasons. First, intentions are not always triggered in straightforward ways by agents' environments or perceptions; to suppose otherwise is to ignore the very phenomena, decision and planning, which make intention so interesting.  Second, motor representations are also not always triggered in any straightforward way by agents' environments or perceptions either.  For these reasons there seems to be no hope of fully explaining matching by postulating common causes for intentions and motor representations.

If common cause explanations are ruled out, another natural approach is to  appeal to content-respecting causal processes. Perhaps, for example, intentions with certain contents (concerning grasping, say) reliably cause motor representations with corresponding contents (also concerning grasping, say). Alternatively we might suppose, very crudely, that some comparator process checks that the contents of motor representations are appropriate given what the agent intends.  Either way, the idea is to explain matches between outcomes specified by the contents of certain states in terms of content-respecting causal processes linking those states.

This type of explanation is arguably appropriate where the states in question have the same representational format.  For example, this type of explanation would arguably be appropriate if our aim were to explain matches between  large-scale intentions and the smaller-scale, more detailed intentions which serve as building blocks for them. But in fact we are concerned with intentions and motor representations which, as argued above, have different representational formats. This creates a potential difficulty.

In general, when two representations differ in format, postulating reliably content-respecting causal processes linking them requires us to explain how their contents are coordinated. To illustrate, suppose you are given some verbal instructions describing a route. You are then shown a representation of a route on a map and asked whether this is the same route that was verbally described. You are not allowed to find out by following the routes or by imagining following them. This puts you in something like the position of the comparator process envisaged above. Special cases aside, answering the question will involve a process of translation because two distinct representational formats are involved, propositional and cartographic. It is not be enough that you could follow either representation of the route. You will also need to be able to translate from at least one representational format into at least one other format. Similarly, for there to be reliably content-respecting causal processes linking intentions with motor representations there would have to be some process of translation.

But why is this a potential difficulty? What is wrong with postulating a process of translation? The difficulty is that nothing at all is known about this hypothetical translation between intention and motor representation, nor about how it might be achieved, nor even about how it might be investigated. Of course this doesn't show that we couldn't fully explain matching by appeal to content-respecting causal processes. But it does show that no such explanation is currently available.

This, then, is why our question about the interface between intentions and motor representations amounts to a problem. It is a problem because of two natural routes to answering the question, the first (appealing to common causes of intentions and motor representations) is a non-starter and the second (appealing to content-respecting causal processes) amounts to no more than a stab in the dark.  Our aim in what follows is to solve the interface problem without postulating either common causes or translation processes.

>>>>

%a4paper or letterpaper (also used in preamble_steve_paper2
\def \papersize {a4paper}

@@

\def \ititle {Shared Intention and Social Aspects of Motor Representation, or How to Avoid Exoticism}
\def \isubtitle {}
\def \iauthor {
S. Butterfill* \& C. Sinigaglia**
\\ 
**Department of Philosophy, University of Warwick
\\ 
***Dipartimento di Filosofia, Università degli Studi di Milano
}
\def \iemail{s.butterfill@warwick.ac.uk}

@@

\begin{abstract}
\noindent
Are there distinctive roles for shared intention and social motor representation in explaining the purposiveness of joint action?  This question needs some background.  A joint action is a purposive action, or something resembling one, comprising two or more agents' actions.  What is the relation between a joint action and the outcome (or outcomes) to which it is directed?  Standard answers to this question appeal to a notion of shared intention and ignore social motor representation.  It is tempting (for some philosophers, at least) to suppose that social motor representations are only enabling conditions for joint action, or else to suppose that social motor representation is merely a variety of shared intention.  The talk will provide reasons for rejecting both suppositions.  The truth may be that shared intention and social motor representation have distinctive roles in explaining the purposiveness of joint action.  This gives rise to a challenge.  On one hand, effective joint action---imagine erecting a tent in a gale together---sometimes requires both shared intentions and social motor representations plus a certain kind of harmony between the two.  On the other hand, recognizing their distinctive roles precludes the existence of direct inferential links between shared intentions and social motor representations.  The challenge is to explain how these two kinds of representation could sometimes harmoniously contribute to effective joint action despite the lack of inferential integration.

\end{abstract}

@@

A purposive joint action is an event involving two or more agents' actions where the actions are collectively directed to some outcome.
What is the relation between a joint action and the outcome or outcomes to which it is directed? 
The standard way of answering this question appeals to shared intention,
a structure of states which stands to joint action roughly as an ordinary, individual intention stands to an ordinary, individual action.  
On the standard view, a joint action is directed to an outcome in virtue of the joint action's being appropriately related to a shared intention which represents this outcome or some related outcome. 
As this view is usually expounded, the relation between joint actions and outcomes to which they are directed is treated as largely independent of the motor processes and representations underpinning action execution. 
Motor representations are usually considered either as philosophically irrelevant enabling conditions, or else as merely filling in additional details of the basic schema provided by the standard story. 
Our aim is to show that this is a mistake:
the standard view must be  refined and extended  to accommodate distinctive roles for intention and motor representation in explaining the purposiveness of joint action.

@@

Twin temptations stand in our way.
The first  temptation is to suppose that motor representation has no bearing at all on our question about the relation between actions and the outcomes to which they are directed. Just as it would be an error to suppose that details of musculoskeletal structure are relevant to this question, so equally---so the temptation---it would be an error to suppose that facts about motor representation are relevant here.

Surrendering to this temptation might be reasonable if all motor representations represented only kinematic or dynamic features of actions, such as mere joint displacements or muscle contractions. However, 
 some motor representations represent action outcomes such as grasping, tearing or throwing.\footnote{ On the notion of action outcome in motor representation, see \citet{jeannerod_motor_2006} and \citet{rizzolatti_mirrors_2008}.}
 Furthermore, as we shall go on to argue, some such representations ground purposive actions. This in outline is why the first temptation should be resisted.

Why accept that there are motor representations of action outcomes? The first step is to consider evidence that motor processes carry information about action outcomes. For any given marker of motor processing (such as a pattern of neuronal discharge or motor-evoked potentials), how can we test whether that marker carries information about action outcomes? The basic principle is straightforward: vary kinematic and dynamic features while holding constant an action outcome; and, conversely, vary action outcomes while holding kinematic and dynamic features constant. 
In order to vary kinematic and dynamic features while holding action outcomes constant, in some studies a single action outcome is achieved using different effectors, hand, mouth or foot, say \citep{rizzolatti:1988_functional,Rizzolatti:2001ug,cattaneo:2010_state-dependent}. A variation on this approach is to contrast performing a grasping action with different tools, so that the same action outcome might require closing or opening the hand depending on the tool used \citep{umilta:2008pliers,cattaneo:2009_representation,rochat:2010_responses}. In order to vary action outcome while holding kinematic and dynamic features constant, researchers have contrasted grasping movements with different distal outcomes such as eating and placing \citep{Fogassi:2005nf,bonini:2010_ventral,cattaneo:2007_impairment}. Another approach is to contrast the same grasping movements performed in the presence or manifest absence of a target object \citep{Umilta:2001zr,villiger:2010_activity}. 
%A related alternative is to contrast the same grasping movements in the presence of objects which could, or manifestly could not, be grasped by means of such movements \citep{koch:2010_resonance}. 
In each of these cases some markers of motor processing are correlated with action outcomes rather than narrowly kinematic or dynamic features of action.\footnote{ Of course some researchers have raised doubts  (e.g.\ \citealp{cavallo_grasping_2011,borroni_mirroring_2011}).  On balance, however, the evidence supports this conclusion.}

So far we have not explicitly distinguished two aspects of action outcomes.  
Action outcomes often specify both a way of acting---whether to grasp or release, say---and also what to act on---on the mug or the pen, say.  
We have been focusing on ways of acting. 
 However a range of behavioural and neurophysiological evidence shows that motor processes also carry information about objects on which actions might be performed.\footnote{
See \citet{buccino:2009_broken,costantini:2010_where,cardellicchio:2011_space,Tucker:1998,tucker:2001_potentiation}.
For a review see \citet{Gallese:2011uq}; for discussion see \citet[pp.\ 410-3]{pacherie:2000_content}. 
}

To say that motor processes involve \emph{information} about action outcomes is not, of course, to say that there are motor \emph{representations} of action outcomes.  To make the step from information to representation we have to show that information about action outcomes guides processing \citep[compare][]{Dretske:1988sq}.

To this end  consider how information about action outcomes is relevant in motor planning, a functional role of motor representation.  Motor planning involves satisfying a variety of requirements. For example, in grasping a mug it is necessary for the hand to prefigure the shape of the mug, to move towards it avoiding potential obstacles and to reach it at a velocity that is both compatible with achieving the type of grip to be used and also suitable given features of the mug such as its fragility and weight \citep{Jeannerod:1995bb,jeannerod:1998nbo}. The need to plan sequences of actions, which may overlap, imposes further requirements. How one should reach for and then grasp a heavy frying pan (say) depends on what one will then do with it. One way of grasping it might be ideal for safely transporting its contents, another for emptying it. 
Requirements on motor planning could not normally be met by explicit practical reasoning, especially given the rapid and fluid transitions involved in many action sequences. 
Rather they require motor processes and representations.
Since 
many of the requirements to be satisfied by motor planning
are partially dependent on action outcomes and not only on  narrowly kinematic or dynamic features of action,
it would be advantageous for some  motor representations to represent action outcomes. 
Given that, as we saw, markers of motor processes carry information about action outcomes, it is now reasonable to conclude that some motor representations are representations of action outcomes.

So far we have argued that motor processes involve representations of action outcomes. It remains to show that such representations ground purposive actions. But this is hardly a further step. How do intentions ground the purposiveness of actions?  On any standard view, an intention represents an outcome, causes an action, and does so in such a way as to facilitate the outcome's occurrence. Similarly, some motor representations  represent action outcomes, play a role in generating actions, and do this in a way that normally facilitates the occurrence of the outcomes represented. To say that motor representations do all this is one way of making precise the metaphor involved in saying that purposive actions are directed to outcomes.

To sum up so far, the first temptation was to suppose that motor representation has no bearing at all on our question about the relation between actions and the outcomes to which they are directed. 
This temptation must be resisted because
motor representations represent action outcomes and ground the directedness of actions to outcomes.  
This fact about motor representation is why a second, complementary temptation is appealing ...

@@

The second temptation is to suppose that motor representation, or some species of it, is a variety of intention, where intention is understood in the standard way as a propositional attitude with a characteristic role in practical reasoning \citep{Bratman:1987xw}. 
Because our claim is that intention and motor representation have distinctive roles in explaining action (§1 on your handout),
we need to show that this temptation should also be resisted.

As background we first need a generic distinction between content and format. Imagine you are in an unfamiliar city and are trying to get to the central station. A stranger offers you two routes. Each route could be represented by a distinct line on a paper map. The difference between the two lines is a difference in content. Each of the routes could alternatively have been represented by a distinct series of instructions written on the same piece of paper; these cartographic and propositional representations differ in format. 
The distinction between content and format is necessary because, as our illustration shows, each can be varied independently of the other.

Our aim now is to show that motor representations differ from intentions with respect to their format. This is why the second temptation must be resisted. That motor representations differ in format from intentions shows that they are genuinely distinct phenomena.

How can we distinguish representational formats? Because a representational format is typically associated with characteristic processing profiles, it is sometimes possible to infer similarities and differences in representational format from similarities and differences in the processes in which representations feature. 
Cognitive neuroscience frequently depends on inferences of just this type. Contrast imagining seeing an object moving with imagining moving an object. 
Imagining seeing and imagining acting involve different processes
 \citep{sirigu:2011_motor,nico:2004_left,vargas:2004s_influence,fourkas:2006_influence}. Importantly for our purposes, imagining seeing an object move and imagining moving an object also differ in ways that indicate differences in representational format. For instance, the way imagining acting unfolds in time is quite different from the way imagining seeing unfolds: how long the former takes is closely related to the time it would take to perform the action, but this is not true of the latter \citep{decety:1989_timing,decety:1996_imagined,Jeannerod:1994oz}. In addition, for object-related actions such as grasping the handle of a cup, manipulating the object in ways that make the action harder (such as orienting the handle to make it less convenient for you to grasp) make a corresponding difference to the effort involved in imagining performing the action \citep{parsons:1994_temporal,frak:2001_orientation}, whereas such manipulations would make no systematic difference to the effort involving in imaging seeing. The explanation for these differences between imagining acting and imagining seeing plausibly involves a difference in the formats of the representations involved.

How do these findings bear on motor representation? Contrast imagining an action with actually performing that action. There is substantial evidence that imagining an action involves many of the same processes which would be involved in performing that action up to the actual muscle contractions \citep{jeannerod:1995_mental,jeannerod:2003_mechanism}. 
There is also independent evidence that the representations underpinning imagining acting have the same format as those involved in actually acting. 
We can conclude, then, that imagining and executing actions are both underpinned by motor representations, and that motor representations differ in format from representations involved in both seeing and imagining seeing.

So far we have been arguing that motor and visual representations differ in format.  
But what we need to show is that motor representations differ in format from intentions.
Why accept this?
We suppose that intentions are propositional. 
This is because of their role in practical reasoning and of the fact that one can have intentions involving quantification and identity; for example, one can intend that one cross seven distinct bridges in 48 hours without yet specifying which bridges or hours.\footnote{ Of course some use the term `intention' for non-propositional representations involved in the execution and control of action.  This is a narrowly terminological issue.} 
But we can also be sure that motor representations differ from propositional representations at least as much as they differ from visual representations given what we have just seen of the characteristic processing profile associated with them.
So no motor representation could be an intention.

@@

We have just argued for three claims. First, some motor representations represent outcomes (rather than, say, only bodily movements). Second, there are actions whose directedness to an outcome is grounded in motor representation. And third, motor representation differs from intention with respect to representational format. A consequence is that a single purposive action may involve representations of the outcomes to which it is directed in at least two different representational formats, motor and propositional. This contributes to a problem we call \emph{the interface problem} ...

Imagine that you are strapped to a spinning wheel facing near certain death as it plunges you into freezing water. To your right you can see a lever and to your left there is a button. 
You form an intention to pull the lever, hoping that this will stop the wheel. If things go well, and if intentions are not mere epiphenomena, this intention will result in your reaching for, grasping and pulling the lever. These actions---reaching, grasping and pulling---may be directed to specific outcomes in virtue of motor representations which guide their execution. It might not be an accident that you both intend to pull a lever and you end up with motor representations of reaching for, grasping and pulling that very lever, so that the outcomes specified by your intention match those specified by motor representations. If this match between outcomes variously specified by intentions and by motor representations is not  accidental, what explains it?

The interface problem is the problem of answering this question, of explaining how intentions and motor representations, with their distinct representational formats, are related in such a way that, in at least some cases, the outcomes they specify non-accidentally match. 
(This question is more precisely stated on your handout.\footnote{Let us say that two collections of outcomes, A and B, \emph{match} in a particular context just if, in that context, either the occurrence of the A-outcomes would normally constitute or cause, at least partially, the occurrence of the B-outcomes or vice versa.
In some cases (i) a particular action is guided both by one or more intentions and by one or more motor representations,
(ii) the outcomes specified by the intentions match the outcomes specified by the motor representations, and
(iii) this match is not accidental. 
The question is, How do non-accidental matches come about?
})

Why think that this question poses a problem at all?  
To see why, consider how the question might be answered.

One might try to explain the match by supposing that intentions and motor representations have a common cause. If the mere presence (or the mere perception) of a lever invariably triggered intentions and motor representations specifying grasping (say), it might be possible to explain matching in this way. But this sort of consideration cannot provide a full explanation of non-accidental matching 
because
intentions are not always triggered in straightforward ways by agents' environments or perceptions.
To suppose otherwise is to ignore the very phenomena, decision and planning, which make intention so interesting.

Another natural approach is to  appeal to content-respecting causal processes. Perhaps, for example, intentions with certain contents (concerning grasping, say) reliably cause motor representations with corresponding contents (also concerning grasping, say). Alternatively we might suppose, very crudely, that some comparator process checks that the contents of motor representations are appropriate given what the agent intends.

There is a difficulty for any  explanation along these lines.
As argued,
intention and motor representation involve different representational formats. 
In general, when any two representations differ in format, postulating reliably content-respecting causal processes linking them requires us to explain how their contents are coordinated. 
To illustrate, suppose you are given some verbal instructions describing a route. You are then shown a representation of a route on a map and asked whether this is the same route that was verbally described. You are not allowed to find out by following the routes or by imagining following them. This puts you in something like the position of the comparator process envisaged above. Special cases aside, answering the question will involve a process of translation because two distinct representational formats are involved, propositional and cartographic. 
Similarly, for there to be reliably content-respecting causal processes linking intentions with motor representations there would have to be some process of translation.

But what is wrong with postulating a process of translation? The difficulty is that nothing at all is known about this hypothetical translation between intention and motor representation, nor about how it might be achieved, nor even about how it might be investigated. Of course this doesn't \emph{show} that we couldn't fully explain matching by appeal to content-respecting causal processes. But it does show that no such explanation is currently available.

This, then, is why our question about the interface between intentions and motor representations amounts to a problem. It is a problem because of two natural routes to answering the question, the first (appealing to common causes of intentions and motor representations) is a non-starter and the second (appealing to content-respecting causal processes) amounts to no more than a stab in the dark.  Our aim for the rest of this talk is to solve the interface problem without postulating either common causes or translation processes.
